Adaptive Sequential Model Selection

### 2014-12-31 paper.tex
added knockoffs and updated paper

\usepackage[round]{natbib}
\usepackage{amsmath,amssymb,amsthm,bm,enumerate,mathrsfs,mathtools}
\usepackage{latexsym,color,verbatim,multirow}
\usepackage{graphicx}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{tikz}
\usepackage[margin=1.3in]{geometry}
\usetikzlibrary{shapes,arrows}
\tikzstyle{block} = [rectangle, draw, fill=white!20,
    text width=7em, text centered, rounded corners, minimum height=4em]
\tikzstyle{title} = [text width=7em, text centered, font=\bfseries]
\tikzstyle{line} = [draw, -latex']

@@

\begin{example}[Linear Regression Variable Selection]\label{ex:linReg}
  Assume that
  \begin{equation}
    Y \sim N(X\beta, \sigma^2I).
  \end{equation}
  With many of the $\beta_j=0$. Our goal is to find a small subset of variables containing the true model.
\end{example}

@@

Suppose that after proposing a random sequence of nested models 
\begin{equation}
  M_0(Y) \sub M_1(Y) \sub \cdots \sub M_d(Y) \sub \cF,
\end{equation}
we wish to sequentially test each $M_k$ against $M_{k-1}$. Let $\bM(Y)$ denote the sequence $(M_0,\ldots,M_d)$, and define its {\em completion index} $K$ to be the index of the first model containing $F$, i.e.
\begin{equation}
  K(\bM) = \min \{k:\, F\in M_k\}.
\end{equation}
Let $N_k$ denote the event $\{Y\in M_{k-1}\}$, i.e. the event that the null hypothesis for the test at stage $k$ is true. $Y\in N_k$ if and only if $k>K$.

For example, suppose that we use the LARS algorithm to build up a linear regression model. Let $J_k(Y)$ be the index of the $k$th variable to have a nonzero coefficient in the LARS path, so that we add variables to the model in the order $X_{J_1}, X_{J_2}, \ldots, X_{J_d}$. It is natural to take model $M_k(Y)$ as
\begin{equation}
  M_k:\, Y \sim 
  N\left(\alpha + X_{J_{1}} \beta_{J_{1}} + \cdots X_{J_k}\beta_{J_k}, 
  \;\sigma^2I\right).
\end{equation}
If the sequence $J_1,\ldots,J_d$ were deterministic instead of adaptive, then we might construct an ANOVA table for testing whether there was sufficient evidence to add each variable to the model in sequence. If the true mean of $Y$ is linear in $X_{J_1}, X_{J_3},$ and $X_{J_5}$, then $K = 5$ and $Y\in N_k$ for $k=6,\ldots,d$.

@@

>>>>

\usepackage[round]{natbib}
\usepackage{amsmath,amssymb,amsthm,bm,enumerate,mathrsfs,mathtools}
\usepackage{latexsym,color,verbatim,multirow}
\usepackage{graphicx}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{tikz}
\usepackage{geometry}
\usetikzlibrary{shapes,arrows}
\tikzstyle{block} = [rectangle, draw, fill=white!20,
    text width=7em, text centered, rounded corners, minimum height=4em]
\tikzstyle{title} = [text width=7em, text centered, font=\bfseries]
\tikzstyle{line} = [draw, -latex']

@@

\section{Introduction}

In many data analysis problems, it is difficult to specify in advance a probabilistic model that is simultaneously expansive enough to adequately capture important characteristics of the sampling distribution, and parsimonious enough to yield precise and interpretable inferences. For example, in a linear regression problem, despite having measured thousands of covariates we commonly expect that only a few are actually relevant. That is, if the design matrix is $X\in \R^{n\times d}$, we expect that for some small subset $M\sub \{1,\ldots,d\}$ of the predictor variables, the model
\begin{equation}
  Y \sim N(X_M\beta^M, \sigma^2I_n)
\end{equation}
approximately holds. In this setting, a common goal is to find a small subset of variables that include $M$.

Analysis of variance (ANOVA) tables provide a classical tool for selecting from among a fixed sequence of nested linear models 
\begin{equation}
  \emptyset = M_0 \sub M_1 \sub \cdots \sub \{1,\ldots,d\},
\end{equation}
using $t$- or $F$-tests to compare each $M_k$ to the next larger proposed model $M_{k+1}$.

In exploratory settings it is commonly impossible to specify a sequence of nested models in advance. Procedures abound for adaptively selecting a sequence of nested models, such as forward stepwise regression, LARS, the lasso, and many more \citep{ESL}. However, there are few accepted methods for choosing a final $M_k$ from among the sequence. In particular, using unadjusted $t$- or $F$-tests is no longer a valid option after adaptive selection. For example, if our algorithm begins by selecting the variable most correlated with $Y$, then the $t$-statistic for testing $M_1$ against $M_0$ is very likely to be large, even if $M_0$, the global null hypothesis, is true.

Cross-validation treats $k$ as a tuning parameter and seeks a value $k$ that will produce a good predictive model. However, it does not yield statements of accuracy or inaccuracy for the models selected.

We propose a modular method for selecting a stopping rule based on the selective framework of \citet{fithian2014optimal}. Our method generalizes the ANOVA table to nested sequences of linear or other probabilistic models that are adaptively selected using $Y$. Given data $Y$, a nested model sequence, and the algorithm used to form the sequence, we return a $p$-values $p_k$ for the test at each step $k$. If $M_K$ is the first correct model (note $K$ is random because the sequence is), the resulting $p$-values $p_{K}, \ldots, p_d$ are independent and (stochastically larger than) uniform, conditional on the selected model sequence. By combining our $p$-values with the sequential testing methods of \citet{wager2014}, we can estimate $K$, and control the probability of selecting an index much bigger than the true $K$.

\WFcomment{Put in real data example?}

\WFcomment{Put in non-Gaussian example?}

@@

Assume that we observe data $Y\in\cY$, with unknown sampling distribution $F$. Our goal is to choose a suitable model for $F$ after using some algorithm to propose an adaptive sequence of nested models:
\begin{equation}
  M_0(Y) \sub M_1(Y) \sub \cdots \sub M_d(Y) \sub \cF.
\end{equation}
Specifically, for $k=1,\ldots,d$, we wish to test whether $M_k$ represents a meaningful improvement on $M_{k-1}$; e.g., test whether adding the $k$th selected variable improves the model with the first $k-1$ variables. Let $\bM(Y)$ denote the sequence $(M_0,\ldots,M_d)$, and define its {\em completion index} $K$ to be the index of the first model containing $F$:
\begin{equation}
  K(\bM) = \min \{k:\, F\in M_k\}.
\end{equation}
Let $N_k$ denote the event $\{F\in M_{k-1}(Y)\}\sub \cY$, the event that the null hypothesis is true for the test at stage $k$. That is, if $M_0$ and $M_1$ are false but $M_2$ is true, then $Y\notin N_0$ or $N_1$, but $Y\in N_2 \sub N_3 \sub \cdots$. Note $Y_k\in N_k$ if and only if $k>K$.

For example, suppose that we use the LARS algorithm to build up a linear regression model. Let $J_k(Y)$ be the index of the $k$th variable to have a nonzero coefficient in the LARS path, so that we add variables to the model in the order $X_{J_1}, X_{J_2}, \ldots, X_{J_d}$. It is natural to take model $M_k(Y)$ as
\begin{equation}
  M_k:\, Y \sim 
  \cN\left(\alpha + X_{J_{1}} \beta_{J_{1}} + \cdots X_{J_k}\beta_{J_k}, 
  \;\sigma^2I\right).
\end{equation}
%If the sequence $J_1,\ldots,J_d$ were deterministic instead of adaptive, then %we might construct an ANOVA table for testing whether there was sufficient %evidence to add each variable to the model in sequence. 
If the true mean of $Y$ is linear in $X_{J_1}, X_{J_3},$ and $X_{J_5}$, then $K = 5$ and $Y\in N_k$ for $k=6,\ldots,d$.

@@

\section{Discussion}

We have seen that sequential selection of models --- the

\bibliographystyle{plainnat}
\bibliography{biblio}
